----
title:    "Sn50 random notes on R"
subtitle: "many while taking stat or epi classes"
date:     '2022'
output:   html_document
----


This is my notebook/scratchpad for R notes.
The basic stuff was in 
[psg/R.html](https://tin6150.github.io/psg/R.html)
but as class uses more advance constructs, and the graphics, 
it is just easier to have Rmd to play with the code in R studio as well.
Knit to html and the result is still visible in the web :)



```{r load library, message = FALSE, warning = FALSE}
library(pacman)
p_load(DT)

```

This example from wk 13 of PHW251 on epicurves by Manasa Susarla:
[PHW251 R for Public Health--Fall 2022 git repo](https://github.com/PHW290/PHW251_Fall2022)


```{r chunk-name-here, include=T, echo=T}

# author: "Manasa Susarla"
# epidemic curvee (wk 13 of PHW251 EpidemicCurves.Rmd)

library(tidyverse)
library(magrittr)
library(pacman)
pacman::p_load(rio)
pacman::p_load(incidence2)

Ebola_data <- import("https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.xlsx")
head(Ebola_data)

##First make sure that your date column is in Date Format
Ebola_data$date_onset <- as.Date(Ebola_data$date_onset)

#Step 1: Create an incidence object
epi_day <- incidence(x = Ebola_data, date_index = date_onset, interval = "day")

#Step 2: Plot
plot(epi_day)


# so much simpler than ggplot !

```


xref:


- [PHW251 R for Public Health--Fall 2022 git repo](https://github.com/PHW290/PHW251_Fall2022)
- [PHW251 project hostec by Github Pages](https://pages.github.com/) and is accessible at :
[https://tin6150.github.io/phw251_group_z/milestone4_groupZ.html](https://tin6150.github.io/phw251_group_z/milestone4_groupZ.html) and
[https://tin6150.github.io/phw251_group_z/milestone5_groupZ.html](https://tin6150.github.io/phw251_group_z/milestone5_groupZ.html)

- <A HREF="https://www.rstudio.com/resources/cheatsheets/">list of R cheatsheets</A>
